\documentclass[letter]{article}
\usepackage{amsmath, amsthm, amssymb, multicol}
\usepackage[margin=0.5in]{geometry}
\usepackage{tikz}

\pagenumbering{gobble}

\begin{document}

\begin{multicols}{2}

	\section{Informal Proofs}
	Informal proofs involve proving stuff with a combination of math and English.\\\\

	\textbf{Proposition}: A less important theorem\\
	\textbf{Lemma}: A proposition used to prove a theorem\\
	\textbf{Corollary}: A theorem that follows directly from another theorem\\
	\textbf{Conjecture}: A unproven statement proposed to be true\\

	\subsection{Methods of Proving}
	\textbf{Direct Proofs:} Prove something of the form $p \rightarrow q$\\
	\textbf{Proof by Contraposition:} Prove $\neg q \rightarrow \neg p$\\
	\textbf{Trivial Proofs:} For $p \rightarrow q$ prove that $q = T$ independent of $p$\\
	\textbf{Proof by Contradiction:} Assume that $p$, then prove that assumption wrong\\
	\textbf{Proof of Equivalence:} Prove $p \leftrightarrow $q by proving $p \rightarrow q$ and $q \rightarrow p$\\
	\textbf{Counterexamples:} Sometimes the easiest way to prove something is wrong is to find an example\\

	\section{Proofs by Induction}

	\section{Number Theory}

	\section{Set Theory}
	$
	\mathbb{N} = \{0,1,2 \ldots \} = \mathbb{Z}_0\\
	\mathbb{R} = \text{real numbers}\\
	\mathbb{Z} = \text{integers}\\
	\mathbb{R}^{+} = \text{positive reals} = \mathbb{R}_{>0}\\
	\mathbb{Z}^{+} = \text{positive ints} = \mathbb{Z}_{>0}\\
	\mathbb{C} = \text{complex numbers}\\
	\mathbb{Q} = \text{rational numbers} = \{\dfrac{p}{q}|p, q\in\mathbb{Z},q\neq 0\}\\
	=\{a+ib|a,b\in \mathbb{R} \text{ and is the imaginary unit.}\}\\
	$
	\textbf{Cardinality}:|A| = the number of elements in A\\
	\textbf{Power set}:P (s) denotes a powerset, which is the set of all
	subsets of s\\
	\textbf{First order logic}:$\forall{a}$ for all options,$\exists{a}$ there
	exists a case.
	\section{Sequences and Series}
	\textbf{Special sequences}\\
	Geometric progression $\{ar^0,ar^1,ar^2,\ldots,ar^n,\ldots\}$\\
	Arithmetic progression $\{a+0d,a+1d,a+2d,\ldots,a+nd,\ldots\}$\\
	\textbf{recurrence relation}: an equation that expresses $a_n$as a function
	of one or more previous terms.
	\section{Combinatorics}
	\subsection{Permutations and Combinations}
	\textbf{r-permutation}: A permutation of a set using r number of
	elements. \\
	\textbf{Number of permutations}: A r-permutation of set with $n$ distinct
	elements.
	$$P(n, r) = \dfrac{n!}{(n - r)!}$$
	\textbf{r-combination}: A subset of the set with $r$ elements. \\
	The number of r-combinations of a set with $n$ elements is the number of
	r-permutations, also known as the binomial coefficient.
	$$C(n, r) = \binom{n}{r} = \dfrac{n!}{(n - r)!r!}$$
	\textbf{Binomial theorem}: $x$ and $y$ are variables and $n$ is a
	non-negative integer. \\
	$${(x + y)}^{n} = \sum{n}{j = 0}\binom{n}{j}x^{n - j}y^{j}$$
	\textbf{Pascal's Identity}: If $n$ and $k$ are integers where
	$n \geq k \geq 0$, then
	$$\binom{n + 1}{k}=\binom{n}{k - 1}+\binom{n}{k}$$
	\textbf{Permutation with reputation}: A string of length $r$ with $n$
	choices $n^r$ If no repetition allowed, then
	$$P(n, r) = \dfrac{n!}{(n - r)!}$$
	\textbf{Combinations with repetition}: The number of r-combinations from a
	set with $n$ elements when repetition of elements is allowed is
	$$C(n + r - 1, r)=C(N + r - 1, n - 1) = \binom{n + r - 1}{r}$$
	\textit{Stars and Bars}: When choosing $6(r)$ cookies from $4(n)$ kinds of
	cookies how many ways to place 3 bars? \\
	\textit{Number of solutions}: How many solutions are there to
	$x_1 + x_2 + x_3 = 11$, where all variables are non-negative integers? \\
	$$$$
	$$\binom{r + n - 1}{n - 1} = \binom{3 - 1 + 11}{3 - 1} = \binom{9}{3}$$
	\textbf{Permutations with indistinguishable Objects}: \\
	\textit{Example}: How many ways can ``SUCCESS'' be re-ordered? \\
	Considering there are 3 S, 2 C, 1 U and 1 E, we can use the template equation
	$$\frac{n!}{n_1!n_2! \ldots n_k!}$$
	The divisor is the number of distinguishable characters. The solution equation
	is:
	$$\frac{7!}{3!2!1!1!}$$
	\textbf{Division rule}: Suppose a task can be carried out $n$ ways, but for
	any way $w$, there are $(d - 1)$ ways that are considered identical to $w$.
	Then, there are $n/d$ ways to do the task. \\
	\textit{How many ways are there to seat four people at a table?} \\
	Two seatings are considered the same when each person has the same neighbors.
	$$\frac{4 \times 3 \times 2 \times 1}{4} = \frac{4!}{6} = 6$$
	\textbf{Ramsey Party Problem}
	\textbf{Ramsey Numbers}: $R(m, n)$ for $m$ and $n$ positive integers greater
	than 1, the minimum number of people at a party such that there are either $m$
	mutual friends of $n$ mutual enemies.
	\subsection{Pigeonhole Principle}
	If $k$ is a positive integer and $k + 1$ objects are placed into $k$ boxes,
	then at least one box contains two or more objects. $N$ pigeons placed into
	$k$ holes would be
	$$\lceil\dfrac{N}{k}\rceil$$

	\section{Discrete Probability}
	\subsection{Laplace's Theory or Probability}
	Given $S$ is a finite set of equally likely outcomes, and $E$ is an event of
	$S$:
	$$\frac{|E|}{|S|} = \frac{\text{Size of } E}{\text{Size of } S}$$
	The probability of $\neg p(E) = 1 - p(E)$. $\neg p(E)$ is the complement of
	$P(E)$. \\
	Given $E_1$ and $E_2$ are events in $S$:
	$$p(E_! \cup E_2) = p(E_1) + p(E_2) - p(E_1 \cap E_2)$$
	\subsection{Probability Distributions}
	Let $S$ be a finite set of outcomes. We can assign a $p(s)$ to each $s$ in $S$
	on two conditions:
	\begin{itemize}
		\item $0 \leq p(s) \leq 1$ for each $s$
		\item $\sum{p(s)} = 1$
	\end{itemize}
	$p$ is a function called the ``\textbf{probability distribution}.'' \\
	If each element in $S$ is equally likely, then the probability of each element
	is $\frac{1}{|S|}$. \\
	The probability of event $E$ is the sum of the probabilities of the outcomes in
	$E$.
	\subsection{Conditional Probability}
	$p(E|F)$ is ``the probability of $E$, given $F$ is true.'' \\
	Given events $E$ and $F$ where $p(E) \neq 0$ and $p(F) \neq 0$:
	$$p(F|E) = \frac{p(E|F)p(F)}{p(E|F)p(F) + p(E|\neg F)p(\neg F)}$$
	\subsection{Bayes' Theorem}
	Given $E$ and $F$ are events and $p(F) \neq 0$:
	$$p(E|F) = \frac{p(F|E)p(E)}{p(E)}$$
	\subsection{Independence}
	Events $E$ and $F$ are \textbf{independent} if and only if
	$$p(E \cap F) = p(E)p(F)$$
	\subsection{Random Variables and Expectation}
	A \textbf{random variable} is a function from the sample space of all real
	numbers that assigns a real number to each possible outcome.
	\textit{Example}: Flip a coin $3$ times. $X(t)$ can equal the number of heads
	in outcome $t$. \\
	The \textbf{probability distribution function} is the probability of the
	random variable $X$ to take a value $r$. \\
	The \textbf{Expected value (expectation)} of a random variable $X$ on sample
	$S$ is
	$$E(X) = \sum p(t)X(t) = \sum p(X = r)r$$
	The \textbf{linearity of expectations} states that given $X_i$, where
	$i \leq 1 \leq n$, where all $X$ are random values in the sample space $S$,
	and $a, b \in \mathbb{R}$, then
	$i = 1, 2, \ldots, n$, then
	$$\sum_{i = 1}^n X_n = \sum_{x = 1}^n E(X_i)$$
	$$E(aX + b) = aE(x) + b$$

	\section{Models of Computation}
	\subsection{Regular Languages}
	\begin{itemize}
		\item A \textbf{languages} is any subset of $\Sigma^*$
		\item Example: $\Sigma = {a, b}$ \\
					${\lambda}$ \\
					${a, aa, aab}$ \\
					${\lambda, a, aa, aaa, aaaa, aaaaa}$ \\
		\item We say a Finite Automata accepts a string if, when it processes that
					string, it ends in an accept state.
	\end{itemize}
	\subsection{Finite Automata and Regular Expressions}
	Example of a finite automata:
	\begin{center}
		\begin{tikzpicture}[scale=0.2]
			\tikzstyle{every node}+=[inner sep=0pt]
			\draw [black] (15.1,-29.5) circle (3);
			\draw (15.1,-29.5) node {$q_0$};
			\draw [black] (24.6,-29.2) circle (3);
			\draw (24.6,-29.2) node {$q_1$};
			\draw [black] (24.6,-29.2) circle (2.4);
			\draw [black] (35.3,-29.2) circle (3);
			\draw (35.3,-29.2) node {$q_2$};
			\draw [black] (13.777,-26.82) arc (234:-54:2.25);
			\draw (15.1,-22.25) node [above] {$a$};
			\fill [black] (16.42,-26.82) -- (17.3,-26.47) -- (16.49,-25.88);
			\draw [black] (18.1,-29.41) -- (21.6,-29.29);
			\fill [black] (21.6,-29.29) -- (20.79,-28.82) -- (20.82,-29.82);
			\draw (19.87,-29.88) node [below] {$b$};
			\draw [black] (27.6,-29.2) -- (32.3,-29.2);
			\fill [black] (32.3,-29.2) -- (31.5,-28.7) -- (31.5,-29.7);
			\draw (29.95,-29.7) node [below] {$a,\mbox{ }b$};
			\draw [black] (33.977,-26.52) arc (234:-54:2.25);
			\draw (35.3,-21.95) node [above] {$a,\mbox{ }b$};
			\fill [black] (36.62,-26.52) -- (37.5,-26.17) -- (36.69,-25.58);
			\draw [black] (7.3,-29.5) -- (12.1,-29.5);
			\fill [black] (12.1,-29.5) -- (11.3,-29) -- (11.3,-30);
		\end{tikzpicture}
	\end{center}
	A machine only accepts \textbf{one} language. If it accept no strings, it
	recognizes the langauge $\o$. \\
	A palindrome is a string whose reversal is identical to the string.
	\subsection{Contex-Free Languages and Grammars}
	\begin{itemize}
		\item A \textbf{regular expression} is a way of describing a regular
					language.
		\item A grammar gives the rules for expressing a language.
		\item A \textbf{context-free} language can be described by a
					\textbf{context-free grammar}.
		\item Theorem: A language is context-free if and only if some
					(nondeterministic) pushdown automaton recognizes it.
		\item A CFG is \textbf{ambiguos} if some string $w \in L(G)$ has at least
					two different leftmost derivations.
		\item Some context-free languages habe only ambiguous grammars.
	\end{itemize}
	Example: Designe a CFG for $L(G) = \{ww^R : w \in \{a, b\}^*\}$ \\
	\begin{align*}
		S & \rightarrow aSa     \\
		S & \rightarrow bSb     \\
		S & \rightarrow \lambda \\
		S & \rightarrow a|b
	\end{align*}
	\subsection{Turning Machines}
	A turing machine:
	\begin{itemize}
		\item Reads input from tape, one character at a time.
		\item Changes state based on input.
		\item Outputs ``accept'' or ``reject'' when input ends.
	\end{itemize}
	In each step, a Turing machine \textbf{reads a symbol, writes a symbol, and
	moves left or right.} A turing machine cannot move left past the start of a
	tape, rather, it stays in place. \\
	A Turing machine has an accept stte and and a reject state. If it reaches one
	of these states, it outputs accept or rejects and stops reading input. An
	infinate loop means ``reject.'' \\

	\section{Limitations of Computational Models}
	\begin{itemize}
		\item The language consisting of all Turing Machines encodings is an
					infinate set.
		\item Not all languages are recognizeable by a Turing Machine.
		\item Infinte sets come in different sizes.
	\end{itemize}
	\subsection{Countable vs. Uncountable Sets}
	\begin{itemize}
		\item An infinite set $S$ is \textbf{countable} if and only if ther exists a
					bijective function $f: S \rightarrow Z^+$. $f$ maps each element of
					$S$ to exactly one element of $Z^+$ and every element of $Z^+$ is
					mapped by some element of $S$, under $f$.
		\item All infinite countable sets are the same size, the same size as $Z^+$.
		\item All finite sets are countable.
		\item \textbf{All subsets of countable sets are countable.}
	\end{itemize}
	\subsection{Enumerators}
	\begin{itemize}
		\item An enumerator for a langage $S$ is a Turing Machine that generates
					all the strings in $S$ one by one. Each string is generated in finite
					time.
		\item \textbf{A set is countable if we can find an enumeration procedure for
					the set.}
		\item Theorem: The set of all Turing Machines is countable.
	\end{itemize}
	\subsection{Languages}
	\begin{itemize}
		\item There are some languages not accepted by Turing Machines.
		\item These languages cannout be described by algorithms.
		\item A languages is called \textbf{Turing recognizable} if some Turing
					Machine recognizes it (aka accepts it).
		\item A language is called \textbf{Turing decidable} if some Turing machine
					decides it.
		\item Not all languages are recognizeable by a turing machine.
	\end{itemize}

	\section{Decision Problems}
	\begin{itemize}
		\item \textbf{Decision problems} are problems with a YES or NO answer.
		\item A language is \textbf{Turing-decidable} if some Turing Machine decides
					it.
		\item The Turing Machine always halts.
		\item Theorem: The membersip problem is undecidable. There is no Turing
					Machine that can decide whether $M$ accepts $w$ for all $M$ and $w$.
		\item Theorem: The halting probelm is undecidable. There is no Turing
					machine that can determine whether $M$ halts on input for $w$ for all
					$M$ and $w$ (Proved in Turing's 1936 paper).
	\end{itemize}

	\section{P and NP}
	$$P \subseteq NP$$
	\begin{itemize}
		\item Every language that can be decided by a deterministic Turing
					Machine in polynomial time can be decided by a nondeterministic
					Turing Machine in polynomial time.
		\item Every deterministic Turing Machine is a nondeterministic Turing
					Machine.
		\item Is $P = NP$? We don't know.
		\item Does the satisfiability problem have a polynomial time deterministic
					algorithm? We don't know.
		\item Cook's Theorem: The satisfiability problem is $NP$-complete.
	\end{itemize}

\end{multicols}

\end{document}
